---
title: "Introduction to Quarto"
format: html
editor: visual
bibliography: references.bib
---

## Quarto

-list -things

**Quarto** enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

Use ctrl + alt + i to quickly add code chunk

Use alt + enter to run the line you are currently on

```{r}
x <- 4 + 8

height_ft <- c(5.2, 6.0, 5.7)

coef <- 3.14

```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

# Things better to do in Visual tab vs Source tab

Insert -\> Citation

[@oke2020]

Table -\> Insert Table

| Col1 | Col2 |
|------|------|
|      |      |
|      |      |

# Exercise

# Introduction

## About the data

-   Water sample data from Arctic Ocean
-   Data downloaded from the Arctic Data Center: [Tweedie ADC dataset](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Setup

See Section 2.3 of textbook for more messages

```{r}
#| message: false

library(readr)
library(here)

# install.packages()
```

## Read in data

```{r}

#| message: false

bg_chem <- read_csv(here::here("data/BGchem2008data.csv"))
```

# Analysis

## Calculate summary statistics

```{r}
colnames(bg_chem)

str(bg_chem)

head(bg_chem)

summary(bg_chem)

unique(bg_chem$Date)

# discussed view()
```

## Calculate mean Redfield ratio

```{r}
# calculated some means

nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)

# chalculate the ratio
ratio <- (nitrate + nitrite + amm ) / phos
```

Here is the ratio we calculated: `r round(ratio)`

## Plot Redfield ratio

```{r}
plot(bg_chem$P, bg_chem$NO3 + bg_chem$NO2 + bg_chem$NH4)
```

# Conclusion
