---
title: "datacleanup"
format: html
editor: visual
---
```{r}

#| message = false


library(dplyr)
library(tidyr)
library(readr)
```



## Data Sources

### Read in the data

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")


```

### Explore data
```{r}
## Prints the column names of my data frame
colnames(catch_original)

## First 6 lines of the data frame
head(catch_original)

## Summary of each column of data
summary(catch_original)

## Prints unique values in a column (in this case, the region)
unique(catch_original$Region)

## Opens data frame in its own tab to see each row and column of the data (do in console)
View(catch_original)
```


## Clean & Reshape data
#### Using the select() function

```{r}
catch_data <- catch_original %>%
    select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

head(catch_data)

catch_data <- catch_original %>%
    select(-All,-notesRegCode)
```

#### Check column types

```{r}
dplyr::glimpse(catch_data)
```

#### Replace values in a column with mutate()

```{r}
catch_clean <- catch_data %>%
    mutate(Chinook = as.numeric(Chinook))

head(catch_clean)

i <- which(is.na(catch_clean$Chinook))
i

catch_data[i,]

catch_clean <- catch_data %>%
    mutate(Chinook = if_else(condition = Chinook == "I", 
                             true = "1", 
                             false = Chinook),
           Chinook = as.numeric(Chinook))

##check
catch_clean[i, ]

```

#### Reshape data with pivot_longer() and pivot_wider()

```{r}
catch_long <- catch_clean %>% 
    #pivot longer all columns except Region and Year
    pivot_longer(
        cols = -c(Region, Year),
        names_to = "species",
        values_to = "catch"
    )

head(catch_long)
```

#### Rename columns with rename()
### Add columns with mutate()
#### Summary statistics using group_by() and summarize()
#### Filtering rows using filter()
#### Sort data with arrange()
#### Split and combine values in columns with separate() and unite()